---
title: "Guidelines"
author: "Florian Mayer, DBCA"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## How the code is written
There are several workflows to write R scripts.

### Scripts creating products
* Data lives on local spreadsheets or on local file shares. (`C:\data\spreasheet.csv`)
* R code lives inside a script. (`script.R`)
* The code runs in the current working directory (bonus is predictable).
* The code is commented inline.
* The script creates local outputs (bonus if in predictable location).

### RMarkdown: story about code
An RMarkdown document where the prose explains the code. 
Code could be data manipulation, data analysis, data visualisation.

#### Lower maturity
Code with inline comments.

#### Higher maturity
Package.

* refactor "the working bits" into pure functions
* document functions
* write test cases
* write vignette with worked examples
* create package from above

### RMarkdown: story about data
An RMarkdown workbook where the code analyses and visualises code, and the prose discusses the findings.

#### Applications
Data ETL workbooks. Examples (DBCA staff only): [Threatened Fauna ETL](https://data.dpaw.wa.gov.au/dataset/79e90d7b-33a3-4aa0-8634-8a12033eb21d/resource/761af921-94ad-4383-be2c-300713c37e98/download/data_etl_fauna.html), 
[Threatened Flora ETL](https://data.dpaw.wa.gov.au/dataset/8075e9ba-1133-446a-8018-4e9b42697746/resource/d1fccebb-02e2-4a40-aba4-c4d6ad012edb/download/data_etl_flora.html).


### RMarkdown: story with data
RMarkdown reports with built-in data analysis and visualisation.
Examples (DBCA staff only):
[Turtle Nesting Census reports](https://data.dpaw.wa.gov.au/dataset/turtle-tracks)

#### Higher maturity
Sweave documents (Latex and R).
Examples: [WAMMP MPA reports](https://data.dpaw.wa.gov.au/dataset/mpa-reports)
Template: [CKAN o' Sweave](https://github.com/dbca-wa/ckan-o-sweave)

## Where the data lives

### Local data
Individual desktop computers

### Intranet
File shares without metadata and clear ownership, no API access

### Data catalogue
Data snapshots with metadata and clear ownership, API access

Example: Get data from data catalogue
```{r data_from_ckan}
```

### Data portals
BioSys, TSC, anything online with an API

Example: Get data from BioSys
```{r data_from_biosys}
```


## Where the products live

### Local products
PNGs in your workdir

### Intranet
File shares

### Data catalogue
Ideal place.

### Cloud storage
For file sharing with external collaborators.

Example: upload to Google Drive.
```{r upload_to_googledrive}
```


## Where the codes lives

### Local code
Individual desktop computers

### Version control
GitHub, with sensitive data excluded

Collaboration through several workflows:

* One master
* forks and pull requests

## Where the code runs

### RStudio Desktop
Per machine

### RStudio Server
Accessible to DBCA staff from public internet.

## How the code runs

### Projects
Keep namespace and environment clean by using projects (from version control, fork this repo or create your own).

### History
Global setting: Never save history
